---
title: "An Interesting Analysis of Tooth Growth"
author: "generalinsight"
date: "July 20, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Overview

Analyzie tooth growth (data:ToothGrowth in R datasets package) to perform basic exploratory analysis, a summary of the data, model to compare tooth growth by **supp** and **dose**, and finally provide conclusions.

### Data Exploration

First, We load the data, and necessary libraries. and perform exploratory analysis.

```{r}
data("ToothGrowth")
library(tidyverse)

str(ToothGrowth)
summary(ToothGrowth)
head(ToothGrowth,10)

```

We plan to model **len** on **supp** (2 level factor) and **dose** (only 3 levels); We change dose to being factors. 


```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
```


### Data Summary

From the plots below, we can see that 


```{r}

ggplot(ToothGrowth, aes(dose, len)) +
        geom_violin(aes(fill = dose)) +
        facet_grid(.~ supp) +
        labs(x = "Dosage", y = "Length", title = "Length vs Dosage by Supp", caption = "Figure - 1: higher dosage (dose) leads to higher tooth length (len), irrespective of supplement (supp)")
        

```


```{r}
ggplot(ToothGrowth, aes(supp, len)) +
        geom_violin(aes(fill = supp)) +
        facet_grid(.~ dose) +
        labs(x = "Supplement", y = "Length", title = "Length by Supplement for Dosage", caption = "Figure - 1: For each dosage (dose), slightly varying tooth lengths (len) per supplement (supp)")
        
```


### Model

Now we move to modeliing the data to thru confidence interval and hypothesis testing 

```{r}

t.test(len ~ supp, ToothGrowth)

```

We find that confidence interval includes 0, and p-value is 0.06 which is greater than 0.05. Therefore, supp ( suplement type) has no impact on tooth growth.

**Below, we are doing a series of comparison on tooth growth by dosage.**

```{r}

t1 <- subset(ToothGrowth, ToothGrowth$dose %in% c(0.5, 1.0))
t.test(len ~ dose, t1)


t1 <- subset(ToothGrowth, ToothGrowth$dose %in% c(0.5, 2.0))
t.test(len ~ dose, t1)

t1 <- subset(ToothGrowth, ToothGrowth$dose %in% c(1.0, 2.0))
t.test(len ~ dose, t1)


```

Based on the results - none of the confidence intervals include 0, as well, the p-values are substanltially small for every tests. 


###Conclusion

Therefore, we can reject null hypothesis, eg we find that tooth length increases with dosage increased. 

This happens irrespective of the different suppplement delivery methods - eg. supp showed to have no effect on tooth growth.







